\documentclass[../en-fa-lab.tex]{subfiles}

\usepackage{hyperref}

\hypersetup{
    pdftitle={(EN) L8 - Disjoint sets},   % The title shown in the browser tab
    pdfauthor={},         % Your name or organization
    pdfsubject={},   % A brief description
    pdfkeywords={}
}

\begin{document}

\section{\texorpdfstring{\textbf{Assignment No. 8: Disjoint Sets}}{Assignment No. 8: Disjoint Sets}}\label{assign8}
\textbf{}

\textbf{Allocated time:} 2 hours

\subsection{Implementation}\label{implementation}

You are required to implement \textbf{correctly} and
\textbf{efficiently} the base operations for \textbf{disjoint set}
(\emph{chapter 21.1} from \cite{cormen})
and the \textbf{Kruskal's algorithm} (searching for the minimum spanning
tree, \emph{chapter 23.2} from \cite{cormen}) using disjoint sets.

You have to use a tree as the representation of a disjoint set. Each
tree holds, besides the necessary information, also the \emph{rank}
field (i.e. the height of the tree).

The base operations on \textbf{disjoints sets} are:

\begin{itemize}
\item
  MAKE\_SET (x)

  \begin{itemize}
  \item
    creates a set with the element \emph{x}
  \end{itemize}
\item
  UNION (x, y)

  \begin{itemize}
  \item
    makes the union between the set that contains the element \emph{x}
    and the set that contains the element \emph{y}
  \item
    the heuristic \emph{union by rank} takes into account the height of
    the two trees so as to make the union
  \item
    the pseudo-code can be found in \emph{chapter
    21.3}\cite{cormen}
  \end{itemize}
\item
  FIND\_SET (x)

  \begin{itemize}
  \item
    searches for the set that contains the element \emph{x}
  \item
    the heuristic \emph{path compression} links all nodes that were
    found on the path to \emph{x} to the root node
  \end{itemize}
\end{itemize}








\subsection{Minimal requirements for grading}\label{minimal-requirements-for-grading}

The lack of any of the minimum requirements (even partially) may result in a lower grade through penalties or refusal to accept the assignment resulting in a grade of 0.

\begin{itemize}
\item
  \emph{Demo}: Prepare a demonstration of correctness for each algorithm implemented. The correctness of each algorithm is demonstrated through a simple example (maximum 10 values).
\item
  The charts created must be easy to evaluate as in grouped and added through the Profiler functions as specified by the assignment requirements. The assignment will not be evaluated if it contains a plethora of ungrouped charts. For example, the comparative analysis implies the grouping of the compared algorithms.
\item
  Interpret the chart and write your observations in the header (block
  comments) section at the beginning of your \emph{main.cpp} file.
\item
  We do not accept assignments without code indentation and with code
  not organized in functions (for example where the entire code is in
  the main function).
\item
  \emph{The points from the requirements correspond to a correct
  and complete solution, quality of interpretation from the block
  comment and \textbf{the correct answer to the questions from the teacher.}}
\end{itemize}












\subsection{Requirements}\label{requirements}

\subsubsection{Correct implementation of MAKE\_SET, UNION and FIND\_SET
(5p)}\label{correct-implementation-of-make_set-union-and-find_set-5p}

\emph{Demo}: The correctness of the algorithm must be proved on a small-sized input.

\begin{itemize}
\item
  create (MAKE) 10 sets + show the contents of the sets
\item
  execute the sequence UNION and FIND\_SET for 5 elements + show the
  contents of the sets
\end{itemize}

\subsubsection{Correct and efficient implementation for Kruskal's algorithm
(2p)}\label{correct-and-efficient-implementation-for-kruskals-algorithm-2p}

\emph{Demo}: The correctness of the algorithm must be proved on a small-sized input.

\begin{itemize}
\item
  create a graph of 5 nodes and 9 edges + \textbf{print the edges}
\item
  apply Kruskal's algorithm + \textbf{print the chosen edges}
\end{itemize}

\subsubsection{Evaluate the disjoint sets operations (MAKE, UNION, FIND)
using Kruskal's algorithm
(3p)}\label{evaluate-the-disjoint-sets-operations-make-union-find-using-kruskals-algorithm-3p}

\textbf{!} Before you start to work on the algorithms evaluation code,
make sure you have a correct algorithm!

Once you are sure your program works correctly:

\begin{itemize}
\item
  vary \emph{n} from 100 to 10000 with a step of 100
\item
  for each n

  \begin{itemize}
  \item
    build an \textbf{undirected}, \textbf{connected}, and
    \textbf{random} graph with random weights on edges (\textbf{n}
    nodes, \textbf{n*4} edges)
  \item
    find the minimum spanning tree using Kruskal's algorithm

    \begin{itemize}
    \item
      evaluate the computational effort of \ul{each individual base
      operation} (MAKE, UNION, FIND\emph{-- resulting in a plot with 3
      series}) on disjoint sets as the sum of the comparisons and
      assignments performed; thus, there should be \textbf{3 series in
      the plot}, one for each operation.
    \end{itemize}
  \end{itemize}
\end{itemize}


\printbibliography

\end{document}